Leadership across cultures and contexts: an exploratory analysis

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Edits in intro cost-benefits section.

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formal institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed). Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
<!-- NEED BETTER CITATIONS ABOVE -->

<!-- BENFITS: MATING -->
Across diverse populations and organizations, community, organization, and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

@@

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved. Leaders therefore are at risk of becoming the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.

>>>>

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formal institutionalized roles and informal ephemeral roles, are thought to be commonly associated with high social status (as previously discussed). Leadership roles are also widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].

<!-- BENFITS: MATING -->
Across diverse populations and organizations, community and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

@@

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. The solutions leaders propose are unlikely to be equally satisfactory for all followers and leaders can become the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
